### 2020-06-16 report/sections/evalconcl/concl/contributions.tex
Report: new conclusion

In this project, we develop and present an
MPST-based framework for developing full-stack
interactive TypeScript applications over WebSocket transport.

We motivate our API generation approach from 
\cite{Hybrid2016,PureScript2019} to generate TypeScript APIs
from a Scribble protocol specification.
Developers write their full-stack applications
by implementing the generated APIs, and can enjoy 
communication safety guarantees by construction.
We explain how our session type encodings target the Node.js
server-side runtime and React.js framework in
\cref{chap:node,chap:react} for server-side
and browser-side endpoints respectively.

In particular, we present a novel approach for integrating
session types into web-based GUI programming based on
translating the theory on model types \cite{MVU2020}
to idiomatic practices on the React.js framework.
With respect to session type theory, implementations using
the generated APIs statically enjoy linear channel usage
guarantees and affine channel usage guarantees for back-end
and front-end targets respectively.

Compared to previous work, we are not only able to
\textbf{statically} provide the same level of communication safety
guarantees from multiparty session type theory, but we do so
using modern web programming practices to increase
the relevance and usability of our work in industry:
our work targets TypeScript, Node.js and React.js,
and we explain in \cref{chap:ext} how we support advanced
web development idioms such as asynchronous implementations.

In addition, existing proposals on API generation for web development
over WebSockets \cite{PureScript2019} only support protocols
implementing server-centric topologies, in order to be 
compatible with WebSocket transport.
We formalise a theory of routed multiparty session types in 
\cref{chap:theory} to prove that it is possible to relax the
server-centric topology assumption over WebSocket transport in a way 
that preserves communication along with the communication safety 
properties inherited from canonical session type theory.

% PLACES
Our work has received positive feedback from academia.
The project has been accepted and published as 
\emph{Generating Interactive WebSocket Applications in TypeScript}
in Volume 314 of the Electronic Proceedings in Theoretical
Computer Science \cite{PLACES2020}; we were also invited to present this work
at the 12th International Workshop on
Programming Language Approaches to Concurrency- \& 
Communication-cEntric Software (PLACES 2020).

>>>>

In this project, we develop and present a novel
MPST-based framework for developing full-stack
interactive TypeScript applications over WebSocket transport.

We motivate our API generation approach from 
\cite{Hybrid2016,PureScript2019} to generate TypeScript APIs
from a Scribble protocol specification.
Developers write their full-stack applications
by implementing the generated APIs, and can enjoy 
communication safety guarantees by construction.
We explain how our session type encodings target the Node.js
server-side runtime and React.js framework in
\cref{chap:node,chap:react} for server-side
and browser-side endpoints respectively.
In particular, we present a novel approach for integrating
session types into web-based GUI programming based on
translating the theory on model types \cite{MVU2020}
to idiomatic practices on the React.js framework.
With respect to session type theory, implementations using
the generated APIs statically enjoy linear channel usage
guarantees and affine channel usage guarantees for back-end
and front-end targets respectively.

Compared to previous work 
\cite{Exceptional,PureScript2019,MVU2020,LINKS} on
session-typed web development, 
we are not only able to
\textbf{statically} provide the same level of communication safety
guarantees from multiparty session type theory, but we do so
using modern web programming practices to increase
the relevance and usability of our work in industry:
our work targets TypeScript, Node.js and React.js,
and we explain in \cref{chap:ext} how we support advanced
web development idioms (such as asynchronous implementations)
and handle premature disconnections gracefully.

We also do not limit support to protocols that describe
a server-centric topology: we formalise 
a theory of routed multiparty session types in 
\cref{chap:theory} to prove that it is possible to relax the
server-centric topology assumption over WebSocket transport in a way 
that preserves communication along with the communication safety 
properties inherited from canonical session type theory.

% PLACES
Our work has received positive feedback from academia.
The project has been accepted and published as 
\emph{Generating Interactive WebSocket Applications in TypeScript}
in Volume 314 of the Electronic Proceedings in Theoretical
Computer Science (EPTCS) \cite{PLACES2020}.
We were also invited to present this work
at the 12th International Workshop on
Programming Language Approaches to Concurrency- \& 
Communication-cEntric Software (PLACES 2020).

### 2020-06-16 report/sections/evalconcl/concl/future.tex
Report: complete chapter 10.1
Report: complete chapter 10 sans code listings for 10.2 and 10.3
Report: new conclusion

\section{Future Work}

@@

\paragraph{Delegation}
Delegation allows channels to be sent between endpoints.
This directly addresses the limitations presented in
\cref{section:evalatm}: in the \tprotocol{ATM} protocol,
when the \trole{Client}
connects to the \trole{Svr},
the \trole{Svr} should be able to send a private channel
for \trole{Client} to perform bank operations,
whilst concurrently accepting new connections.

The API generation solution 
presented by Scalas et al. \cite{LinearDecomp} supports
\textit{distributed multiparty delegation} in Scala.
We have also shown in \cref{section:evalatm} that
the developer can implement delegation with small changes
to the session runtime.
Future work can apply the findings from \cite{LinearDecomp},
extend the Scribble protocol language for sending
channels through messages, and formalise the workaround
in \cref{section:evalatm} with respect to the extended 
Scribble syntax for delegation.

\paragraph{``Serverless'' Server-Side Endpoints}
\dots

>>>>

\section{Future Work}
\label{section:future}

@@

\paragraph{Session Delegation}
Delegation allows channels to be sent between endpoints.
Using the \tprotocol{Noughts and Crosses} game
from \cref{section:evalgame}, the developer
may want the \trole{Svr} to send a private channel
to both players to play the game,
whilst concurrently accepting new connections.

The API generation solution 
presented by Scalas et al. \cite{LinearDecomp} supports
\textit{distributed multiparty delegation} in Scala.
Their work also uses the Scribble toolchain
for protocol specification.
Future work can apply the findings from \cite{LinearDecomp}
and extend \codegen to support channel delegation
in the generated handler-style APIs.

### 2020-06-16 report/sections/evalconcl/eval.tex
Report: progress on chapter 10
Report: new conclusion

To evaluate the expressiveness
of our work,
we use three case studies
of protocols found in webservices
to demonstrate strengths and weaknesses
of our work:

@@

\item
\textbf{\tprotocol{ATM} Protocol:}

In \cref{section:evalatm},
we implement an ATM web application to demonstrate
limitations of our work. Our generated APIs do not allow
the server endpoint (in this case, the \trole{Bank})
to participate in concurrent sessions (i.e. interacting
with different \trole{Client}s concurrently), even though
the protocol is written in a way that allows this.
However, we propose a way to address this limitation, and expand
upon this in \cref{section:future}.

@@

\input{sections/evalconcl/eval/atm}

>>>>

To evaluate the expressiveness
of our work,
we use two case studies
of protocols found in webservices
to demonstrate strengths and weaknesses
of our work:

@@

@@

%\input{sections/evalconcl/eval/atm}
